A universal of human social cognition: Children from 17 communities process gaze in similar ways

### 2024-10-16 paper/paper.tex
long version first draft

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{keywords\newline\indent Word count: X}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Unity in diversity: Children from 17 diverse communities process gaze in similar ways},
  pdfauthor={Manuel Bohn1,2, Julia Prein1,2, \ldots\ldots, \& Daniel B. M. Haun2},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Unity in diversity: Children from 17 diverse communities process gaze in similar ways}
\author{Manuel Bohn\textsuperscript{1,2}, Julia Prein\textsuperscript{1,2}, \ldots{}\textsuperscript{\ldots{}}, \& Daniel B. M. Haun\textsuperscript{2}}
\date{}

\shorttitle{Gaze-following across 17 communities}

@@

Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

Enter author note here.

The authors made the following contributions. Manuel Bohn: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Julia Prein: Conceptualization, Methodology, Writing - Review \& Editing; \ldots{}: Conceptualization, Methodology, Writing - Review \& Editing; Daniel B. M. Haun: Conceptualization, Writing - Review \& Editing.

Correspondence concerning this article should be addressed to Manuel Bohn, Postal address. E-mail: \href{mailto:my@email.com}{\nolinkurl{my@email.com}}

@@

\affiliation{\vspace{0.5cm}\textsuperscript{1} Institute of Psychology in Education, Leuphana University Lüneburg\\\textsuperscript{2} Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology}

\abstract{%
Understanding gaze is critical for human social interaction. In a comprehensive cross-cultural study spanning five continents and 17 distinct cultural contexts, we examined the development of gaze understanding among preschoolers using a single reliable task tailored to each cultural setting. Our data provides evidence that children worldwide process gaze in the same way. Key performance signatures of a computational model that sees gaze following as a form of social vector estimation were found in all communities. Additionally, we found a subtle yet cross-culturally consistent relationship between children's environments and their ability to understand gaze. These results highlight the cross-cultural robustness of core social cognitive skills, as well as similarities in the developmental process, suggesting that fundamental aspects of social cognition and interaction emerge and develop in comparable ways globally and hence are a bedrock of human social cognition.
}

@@

Social cognition is a defining aspect of the human species (\emph{1}--\emph{3}). It is supposed to enable unique forms of communication and cooperation that underlie cumulative cultural evolution and the formation of complex societies (\emph{4}--\emph{6}). The eyes are the proverbial window to the mind and the starting point for a majority of social reasoning processes (\emph{7}). Gaze is used to infer the focus of visual attention, which is a critical aspect of coordinated activities, including communication and cooperation (\emph{8}, \emph{9}). The ability to follow gaze emerges early in childhood (\emph{10}--\emph{12}) and individual differences in children's gaze following ability predict later life outcomes, most notably, later communicative abilities (\emph{13}, \emph{14}). Underlying this narrative is the wide-held assumption that gaze following is fundamental to human social cognition and, therefore, works and develops in the same way across human societies despite substantial variation in developmental contexts. This claim, however, lacks a solid empirical foundation.

We conducted a large-scale cross-cultural study on the development of gaze-following abilities to study potentially universal processing signatures and their development. Previous developmental studies focused on the onset of gaze following in infancy (\emph{15}, \emph{16}). The 1311 participants in the study lived in 17 different communities spread over five continents (Fig. \ref{fig:fig1}A, Tab. \ref{tab:tab1}). The countries from which data was contributed to the study represented 46 \% of the world's population. Communities covered a broad spectrum of geographical locations, social and political systems, languages, and subsistence styles (see Supplemental Materials).

@@

{\centering \includegraphics[width=0.75\linewidth]{../visuals/fig1}

@@

\caption{(A) Data collection sites. Points show approximate geographical location of the data collection sites, coloring shows the sample sizes. (B) Screenshot from the task. The scene depicts the choice phase in a test trial. Participants had to use the gaze of the agent to locate the balloon and click on the hedge where they thought the balloon was. Agents, audio recordings and backgrounds were adpated to each cultural setting. (C) Graphical illustration of the cognitive model. Individuals infer the target of an agent's attention by estimating a vector based on the position of the pupils within the eyes. This process is noisy, illustrated by the different vectors. Individuals differ in their level of precision (indicated by alpha). Regardless of the level of precision, the further out the target lands, the less precise the model predicts individuals to be.}\label{fig:fig1}
\end{figure}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{unicode-math}
\usepackage{setspace}
\usepackage{libertine}
\captionsetup[figure]{font={stretch=1}}
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={A universal of human social cognition: Children from 17 communities process gaze in similar ways},
  pdfauthor={Manuel Bohn1,2,*, Julia Christin Prein1,2,*, Agnes Ayikoru3, Florian M. Bednarski4, Ardain Dzabatou5, Michael C. Frank6, Annette M. E. Henderson4, Joan Isabella3, Josefine Kalbitz2, Patricia Kanngiesser7, Dilara Keşşafoğlu8, Bahar Köymen9, Maira V. Manrique-Hernandez2, Shirley Magazi10, Lizbeth Mújica-Manrique2, Julia Ohlendorf2, Damilola Olaoba2, Wesley R. Pieters10, Sarah Pope-Caldwell2, Katie Slocombe11, Robert Z. Sparks6, Jahnavi Sunderarajan2, Wilson Vieira2, Zhen Zhang12, Yufei Zong12, Roman Stengelin2,10,+, \& Daniel B. M. Haun2,+},
  pdflang={en-EN},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{A universal of human social cognition: Children from 17 communities process gaze in similar ways}
\author{Manuel Bohn\textsuperscript{1,2,*}, Julia Christin Prein\textsuperscript{1,2,*}, Agnes Ayikoru\textsuperscript{3}, Florian M. Bednarski\textsuperscript{4}, Ardain Dzabatou\textsuperscript{5}, Michael C. Frank\textsuperscript{6}, Annette M. E. Henderson\textsuperscript{4}, Joan Isabella\textsuperscript{3}, Josefine Kalbitz\textsuperscript{2}, Patricia Kanngiesser\textsuperscript{7}, Dilara Keşşafoğlu\textsuperscript{8}, Bahar Köymen\textsuperscript{9}, Maira V. Manrique-Hernandez\textsuperscript{2}, Shirley Magazi\textsuperscript{10}, Lizbeth Mújica-Manrique\textsuperscript{2}, Julia Ohlendorf\textsuperscript{2}, Damilola Olaoba\textsuperscript{2}, Wesley R. Pieters\textsuperscript{10}, Sarah Pope-Caldwell\textsuperscript{2}, Katie Slocombe\textsuperscript{11}, Robert Z. Sparks\textsuperscript{6}, Jahnavi Sunderarajan\textsuperscript{2}, Wilson Vieira\textsuperscript{2}, Zhen Zhang\textsuperscript{12}, Yufei Zong\textsuperscript{12}, Roman Stengelin\textsuperscript{2,10,+}, \& Daniel B. M. Haun\textsuperscript{2,+}}
\date{}

\shorttitle{Gaze following across 17 communities}

@@

The authors would like to thank Luke Maurits for statistical advice. Manuel Bohn was supported by a Jacobs Foundation Research Fellowship (2022-1484-00). We are grateful to thank all children and caregivers for participating in the study. We thank the Max Planck Society for the Advancement of Science.

The authors made the following contributions. Manuel Bohn: Conceptualization, Methodology, Formal Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Julia Christin Prein: Conceptualization, Methodology, Software, Investigation, Writing - Review \& Editing; Agnes Ayikoru: Investigation; Florian M. Bednarski: Investigation, Writing - Review \& Editing; Ardain Dzabatou: Investigation; Michael C. Frank: Investigation, Writing - Review \& Editing; Annette M. E. Henderson: Investigation, Writing - Review \& Editing; Joan Isabella: Investigation; Josefine Kalbitz: Investigation, Writing - Review \& Editing; Patricia Kanngiesser: Investigation, Writing - Review \& Editing; Dilara Keşşafoğlu: Investigation, Writing - Review \& Editing; Bahar Köymen: Investigation, Writing - Review \& Editing; Maira V. Manrique-Hernandez: Investigation; Shirley Magazi: Investigation; Lizbeth Mújica-Manrique: Investigation, Writing - Review \& Editing; Julia Ohlendorf: Investigation; Damilola Olaoba: Investigation; Wesley R. Pieters: Investigation, Writing - Review \& Editing; Sarah Pope-Caldwell: Investigation; Katie Slocombe: Investigation, Writing - Review \& Editing; Robert Z. Sparks: Investigation; Jahnavi Sunderarajan: Investigation; Wilson Vieira: Investigation; Zhen Zhang: Investigation, Writing - Review \& Editing; Yufei Zong: Investigation; Roman Stengelin: Conceptualization, Methodology, Investigation, Writing - Review \& Editing; Daniel B. M. Haun: Conceptualization, Funding acquisition, Writing - Review \& Editing.

Correspondence concerning this article should be addressed to Manuel Bohn, Universitätsallee 1, 21335 Lüneburg, Germany. E-mail: \href{mailto:manuel.bohn@leuphana.de}{\nolinkurl{manuel.bohn@leuphana.de}}

@@

\affiliation{\vspace{0.5cm}\textsuperscript{1} Institute of Psychology in Education, Leuphana University Lüneburg\\\textsuperscript{2} Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology\\\textsuperscript{3} Budongo Conservation Field Station\\\textsuperscript{4} School of Psychology, University of Auckland\\\textsuperscript{5} Université Marien Ngouabi\\\textsuperscript{6} Department of Psychology, Stanford University\\\textsuperscript{7} School of Psychology, University of Plymouth\\\textsuperscript{8} Department of Psychology, Koç University\\\textsuperscript{9} Division of Psychology, Communication, and Human Neuroscience, University of Manchester\\\textsuperscript{10} Department of Psychology and Social Work, University of Namibia\\\textsuperscript{11} Department of Psychology, University of York\\\textsuperscript{12} CAS Key Laboratory of Behavioral Science, Institute of Psychology, Chinese Academy of Sciences\\\textsuperscript{*} joint first author\\\textsuperscript{+} joint last author}

\abstract{%
Theoretical accounts assume that key features of human social cognition are universal. Here we focus on gaze following, the bedrock of social interactions and coordinated activities, to test this claim. In a comprehensive cross-cultural study spanning five continents and 17 distinct cultural communities, we examined the development of gaze following in early childhood. We identified key processing signatures through a computational model that assumes that participants follow an individual's gaze by estimating a vector emanating from the eye center through the pupil. We found these signatures in all communities, suggesting that children worldwide processed gaze in highly similar ways. Absolute differences between groups were accounted for by a cross-culturally consistent relationship between children's exposure to touchscreens and their performance in the task. These results provide strong evidence for a universal process underlying a foundational socio-cognitive ability in humans that can be reliably inferred even in the presence of cultural variation in overt behavior.
}

@@

\begin{verbatim}
## # A tibble: 2 x 3
##   side  left_bias right_bias
##   <chr>     <dbl>      <dbl>
## 1 left      0.170      0.830
## 2 right     0.754      0.246
\end{verbatim}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Human socio-cognitive skills enable unique forms of communication and cooperation that provide a bedrock for cumulative culture and the formation of complex societies (Henrich, 2016; Heyes, 2018; Laland \& Seed, 2021; Legare, 2019; Tomasello, 2020; Tomasello \& Rakoczy, 2003; Wellman, 2014). The eyes are the proverbial ``window to the mind'' and eye gaze is essential for many social reasoning processes (Doherty, 2006; Emery, 2000; Frischen et al., 2007; Shepherd, 2010). Others' eye gaze is used to infer their focus of visual attention, which is a critical aspect of coordinated activities, including communication and cooperation (Langton et al., 2000; Richardson \& Dale, 2005; Rossano, 2012; Scaife \& Bruner, 1975; Sebanz et al., 2006; Tomasello et al., 2007). During ontogeny, gaze following is an important aspect of many critical learning objectives that enable children to become functioning members of the society they grow up in, including language, social learning and joint action (Brooks \& Meltzoff, 2005; Brownell, 2011; Carpenter et al., 1998; Moore, 2008; Mundy \& Newell, 2007; Stephenson et al., 2021). Because of the central role gaze following plays during human ontogeny, it has been widely argued that gaze following has been a target of natural selection (Clark et al., 2023; Emery, 2000; Kano, 2023; Tomasello et al., 2007). This implies that the process by which humans use gaze direction to infer the focus of attention is universal. In this paper, we report a comprehensive cross-cultural study on the ontogeny of gaze following in which we shed light on the universal aspects of gaze following as well as sources of variation and their origins.

\hypertarget{ontogeny-of-gaze-following}{%
\subsection{Ontogeny of gaze following}\label{ontogeny-of-gaze-following}}

The ability to follow gaze emerges early in development (Byers-Heinlein et al., 2021; Del Bianco et al., 2019; Gredebäck et al., 2010; Tang et al., 2023). The earliest signs of gaze following have been found in infants as young as four months (Astor et al., 2021; D'Entremont et al., 1997). Throughout the first two years of life, children refine their abilities: they interpret gaze in mentalistic terms (Butterworth \& Jarrett, 1991; Deák et al., 2000), for example, they follow gaze to locations outside their own visual field by moving around barriers (Moll \& Tomasello, 2004). Initially, children rely more on head direction than actual gaze direction (Michel et al., 2021). In fact, when head and gaze direction diverge, children fail to accurately locate the agent's focus of attention up until 19 months of age (Lempers, 1979).

From an evolutionary perspective, while many species are able to follow gaze based on head directions, uniquely human forms of joint action and communication require a more precise localization of other's attention and thus critically rely on gaze direction inferred from eye movements (Emery, 2000; Hessels, 2020). In a recent study, Prein, Maurits, et al. (2024) studied the development of gaze following based on eye movements across from three years up until old age. They found particularly steep developmental improvements in the preschool years resulting in a relatively stable level of accuracy from ten years onward and a slight decrease starting around age 40.

The studies reported thus far, all relied on data collected in western affluent settings. Such settings represent only a minority of the worlds population and are thus insufficient to make claims about universal aspects of human cognition (Amir \& McAuliffe, 2020; Nielsen et al., 2017; Norenzayan \& Heine, 2005). Three studies with infants and children from traditionally underrepresented parts of the world (Bhutan, India, Peru, Vanuatu) find that children start gaze following (including head direction) at similar ages, the rates of gaze following, however, differed between cultural settings (Astor et al., 2022; Callaghan et al., 2011; Hernik \& Broesch, 2019).

Rates and accuracy of gaze following do not just differ between cultural settings; there is also substantial variation within settings. In fact, the pivotal role of gaze following in many uniquely human activities has been studied by relating individual differences in gaze following to other phenomena -- both cross-sectionally and longitudinally (Brooks \& Meltzoff, 2015; Carpenter et al., 1998). For example, gaze following at 10 months predicts language scores at 18 months of age (Brooks \& Meltzoff, 2005; see also Macdonald \& Tatler, 2013). Furthermore, difficulties with gaze following have been linked to developmental disorders, including Autism (Itier \& Batty, 2009; Thorup et al., 2016, 2018) and -- at least in some cultural contexts -- to maternal postpartum depression {[}astor2022maternal{]}. Individual differences are also key to explaining the -- mainly social-interactional -- driving forces behind the development of gaze following. For example, it has been found that early attachment quality or the use of gaze in communicative interactions predict later rates of gaze following (Astor et al., 2020; Movellan \& Watson, 2002; Senju et al., 2015).

\hypertarget{cognitive-universals-and-sources-of-variation}{%
\subsection{Cognitive universals and sources of variation}\label{cognitive-universals-and-sources-of-variation}}

The existence of variation in gaze following both within and between cultural settings raises the question of how to square these findings with the suggestion that gaze following is a fundamental building block of human social cognition and interaction and that eye movements are processed the same way in humans all over the world. Looking at other aspects of social cognition, one could easily make the argument that variation is the norm rather than the exception (Dixson et al., 2018; Mayer \& Träuble, 2013; see e.g., Miller et al., 2018; Taumoepeau et al., 2019; Wellman, 2014). As a first step, answering this question requires data from many different cultural settings. In a second step, however, we need to a way of detecting universal processes in such data. The traditional approach is to compare some sort of aggregate measure (mean level of performance, average age of onset) across cultural settings. Absolute differences in mean performance across communities are interpreted as a signal of different underlying cognitive processes (Blake et al., 2015; see e.g., House et al., 2020; Kanngiesser et al., 2022; Van Leeuwen et al., 2018). No differences seemingly support the existence of a psychological universal. Such an approach, however, neglects the existence of within-cultural variation altogether (see also Gurven, 2018).

In the present study, we want to take a different approach for which we assume that universal processes and variation can co-exist (Greenfield et al., 2003; Jensen, 2012; Kline et al., 2018). Instead of starting with the outcome (performance in the task), we start with the process that generates the outcome. By defining this process, we make a proposal for the universal aspect of the process. At the same time, we define variable aspects in the process that generate individual differences. This allows us to define signatures that the process leaves behind that can be detected independent of absolute levels of performance.

For gaze following, we can use the computational model proposed by Prein, Maurits, et al. (2024) to derive such predictions. They formalized the widely-held view that gaze following involves estimating a vector emanating from the eye center through the pupil (Butterworth \& Jarrett, 1991; Symons et al., 2004; Todorović, 2006; Yaniv \& Shatz, 1990). The key innovation of the model is that it explains how individuals may use the same process but still differ in their measured abilities. The process always involves estimating a vector but also involves a degree of uncertainty because the eye center is not directly observable. Individuals are assumed to differ in their level of uncertainty with which they estimate the vector which causes differences in their observable behavior. Importantly, the assumed process leaves a key signature in the data that is observable independent of the absolute level of performance. In the present study, we therefore focus on this signature instead of absolute levels of performance when evaluating the claim whether there is evidence for a universal cognitive process underlying gaze following.

\hypertarget{the-current-study}{%
\subsection{The current study}\label{the-current-study}}

The present study had three goals. First, to collect a comprehensive data set and study the ontogeny of gaze following beyond infancy across cultures. To make this possible, we used a semi-standardized task that required minimal assistance from an experimenter and no behavioral coding. The task is an animated picture book presented on a tablet screen. Children watched a balloon disappear behind a hedge. An agent followed the trajectory of the balloon with their eyes (Figure \ref{fig:fig1}B). The key dependent variable was (im)precision, that is, the deviation between where the agent looked (where the balloon was) and the child's response. The task's flexible implementation as a browser-based web-app allowed us to quickly tailor its visual and audio content to each cultural setting (visuals and audio). The task has been psychometrically evaluated and has shown to yield reliable individual-level measurements across communities and across ages (Prein, Kalinke, et al., 2024; Prein, Bednarski, et al., 2024).

We collected data in 17 different communities across 14 countries and five continents. Communities covered a broad spectrum of geographical locations, social and political systems, languages, and subsistence styles. This diversity allowed us to overcome a common pitfall of cross-cultural studies that compare urban communities from the Global North to rural communities from the Global South (Barrett, 2020). We aimed for large sample sizes within each community to contrast within- and between cultural variation. Our expectation regarding the first goal was to see substantial variation across cultures but even more variation between individuals. In all communities, we expected performance to increase with age.

The second goals was to look for signatures in the data of the universal gaze following process predicted by the model of Prein, Bednarski, et al. (2024). In the task, the hidden object is located on a horizontal plane at the lower end of the screen. The agent is located in the upper center of the screen (see Figure \ref{fig:fig1}B). The model predicts trials in which the object is hidden further away from the center to be more difficult resulting in higher imprecision. The signature is thus a u-shaped relation between object location and imprecision (Figure \ref{fig:fig12}).

Finally, we sought to explain individual differences in gaze following precision by linking them to methodological aspects of the study as well as aggregate measures of children's everyday social experience. Experience with tablets and touch screens co-varied with community and we expected children more familiar with this medium to perform better. Previous work suggested that gaze following is refined in social interaction (Movellan \& Watson, 2002; Senju et al., 2015). To approximate social interaction, we asked parents to fill out a questionnaire about household size and composition. We acknowledge that this measure approximates opportunities for social interaction in a rather crude way but we nevertheless expected children living in larger households and with more siblings (relative to their community) to be more accurate when following gaze.

@@

{\centering \includegraphics[width=1\linewidth]{../figures/fig1_2}

@@

\caption{(A) Data collection sites. Points show the approximate geographical location of the data collection sites, coloring shows the sample sizes. (B) Screenshots from the task. The upper scene depicts the start, and the lower scene depicts the choice phase in a test trial. Participants had to use the gaze of the agent to locate the balloon and touch the location on the hedge where they thought the balloon was. Agents, audio recordings and backgrounds were adapted to each community. (C) Drawings used as agents across communities.}\label{fig:fig1}
\end{figure}

\begin{center}
\begin{ThreePartTable}

\begin{TableNotes}[para]
\normalsize{\textit{Note.} 1 Proportion of participants who have access to touchscreens according to parental questionnaire. 2 Local collaborators and piloting suggested that Nigerian English is suitable for Windhoek as well.}
\end{TableNotes}

\begin{longtable}{m{0.125\linewidth}m{0.125\linewidth}m{0.125\linewidth}m{0.125\linewidth}m{0.125\linewidth}m{0.125\linewidth}m{0.125\linewidth}}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:tab1}Participant demographics.}\\
\toprule
Continent & Country & Community & N (male) & Age (range) & Language & Touchscreen exposure1\\
\midrule
\endfirsthead
\caption*{\normalfont{Table \ref{tab:tab1} continued}}\\
\toprule
Continent & Country & Community & N (male) & Age (range) & Language & Touchscreen exposure1\\
\midrule
\endhead
Americas & Argentina & Buenos Aires & 105 (53) & 4.72 (3.00 - 6.96) & Spanish (Rioplatense) & 0.90\\
 & México & Ocuilan & 127 (63) & 4.96 (2.57 - 6.95) & Spanish (Mexican) & 0.77\\
 & USA & Stanford & 98 (54) & 4.99 (2.52 - 7.90) & English (American) & 0.98\\
Africa & Namibia & Hai||om & 60 (38) & 5.85 (2.74 - 8.34) & Hai||om & 0.05\\
 &  & Khwe & 59 (24) & 5.84 (3.38 - 8.63) & Khwedam & 0.19\\
 &  & Windhoek & 39 (17) & 5.69 (2.66 - 8.66) & English (Nigerian)2 & 0.95\\
 & Nigeria & Akure & 114 (54) & 5.07 (2.57 - 7.33) & English (Nigerian) & 0.91\\
 & Rep. Congo & BaYaka & 29 (13) & 7.80 (3.94 - 10.56) & Yaka & 0.00\\
 &  & Bandongo & 30 (11) & 7.45 (3.50 - 10.95) & Lingala & 0.00\\
 & Uganda & Nyabyeya & 125 (62) & 5.94 (2.67 - 8.92) & Kiswahili & 0.34\\
 & Zambia & Chimfunshi & 22 (5) & 5.98 (2.88 - 8.00) & Bemba & 0.14\\
Europe & Germany & Leipzig & 100 (48) & 4.88 (2.53 - 6.95) & German & 0.89\\
 & UK & Plymouth & 70 (30) & 6.02 (2.38 - 8.94) & English (British) & 0.99\\
Asia & China & Beijing & 123 (62) & 5.47 (2.69 - 8.48) & Mandarin & 0.95\\
 & India & Pune & 148 (73) & 6.14 (3.06 - 8.83) & English (Indian) / Marathi & 0.93\\
 & Türkiye & Malatya & 85 (40) & 5.02 (2.75 - 7.12) & Turkish & 1.00\\
Oceania & New Zealand & Auckland & 43 (19) & 5.14 (2.81 - 8.75) & English (New Zealand) & 0.95\\
\bottomrule
\addlinespace
\insertTableNotes
\end{longtable}

\end{ThreePartTable}
\end{center}

We used an animated picture book tablet task in which participants had to locate a hidden object based on observing an agent's gaze. Children watched a balloon disappear behind a hedge. An agent followed the trajectory of the balloon with their eyes (Fig. \ref{fig:fig1}B). The key dependent variable was the (im)precision with which children located the agent's focus of attention, that is, the deviation between where the agent looked (where the balloon was) and the child's response. We adapted visuals and audio instructions specifically for each of the 17 communities. Previous work demonstrated excellent individual-level measurement properties for this task in a German sample (Prein, Kalinke, et al., 2024).

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{preregistration}{%
\subsection{Preregistration}\label{preregistration}}
